---
title: "tidy-model"
author: "Makayla Taladoc"
date: "2025-02-17"
output: html_document
---

install.packages(c('tidyverse', 'modeldata'))

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse') ; library('modeldata')
```

## Introduction

This R Markdown document demonstrates my abilities to use models for data analysis using a data set collected on crickets.

```{r load data, include = FALSE}
data(crickets, package = "modeldata")
names(crickets)
```

## What is that sound?

In this report, we examine what factors predict a cricket's chirp rate.

```{r summary, echo = FALSE}
summary(crickets)
ggplot(crickets, aes(x = rate)) +
  geom_histogram(bins = 8) +
  ggtitle("Distribution of the chirp rate of crickets") +
  xlab('Chirp rate (per min.)')
```

## Temperature affects chirp rate

```{r temp, echo = FALSE}
ggplot(crickets, aes(x = temp, y = rate)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')

temp_lm <- lm(rate ~ temp, crickets)

summary.lm(temp_lm)
```

Based on a scatter plot of temperature and chirping and a correlation test, it seems that as temperature increases one degree, the rate of chirping increases about 4.2 chirps per minute.

## Species-specific effects of temperature on chirping

```{r species, echo = FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate for two species of crickets") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')

species_lm <- lm(rate ~ temp + species, crickets)

summary.lm(species_lm)
```
Based on the scatter plot of temperature and chirp rate, the following observations were found:

- For the exclamationis species, as temperature increases by one, the rate of chirping increases by about 3.6 chirps per minute. 

- For the niveus species, as temperature increases by one, the rate of chirping increases by about 37 chirps per minute.

## Interactions 

```{r species histogram, echo = FALSE}
ggplot(crickets, aes(x = rate, fill = species)) + 
  geom_histogram(position = 'identity', alpha = 0.7, bins = 8) + 
  ggtitle("Distribution of the chirp rate of crickets") + 
  xlab('Chirp rate (per min.)')
```

```{r interactions, echo = FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate for two species of crickets") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')

species_x_temp_lm <- lm(rate ~ temp + species + temp:species, crickets)
summary.lm(species_x_temp_lm)

anova(species_lm, species_x_temp_lm)
```

## Setup for diamonds data

install.packages("ggplot2")

library(ggplot2)

diamonds

ggplot2::diamonds

## Worth of a diamond

```{r worth, echo = FALSE}
ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
  geom_smooth(method = "lm") +
  ggtitle("Plot of Predicted Diamond Prices by cut") +
  ylab('Price of Diamond') +
  xlab('Carat of Diamond')

price_lm <- lm(price ~ carat + cut, diamonds)

summary.lm(price_lm)
  
```



Based on the linear regression plot of price of diamond, carat of diamond, and cut, ideal cut diamonds are priced highest in comparison with fair, good, very good, and premium cuts. However, ideal cuts do not extend past a carat of 3.5, thus resulting in the very good cut to be the most expensive diamonds as the carat increases.



